---
title: "Analyzing Age/Length structures in IC"
author: "Youen Vermard"
date: "`r date()`"
output:
  pdf_document: default
  html_document:
    df_print: paged
  word_document: default
---


```{r echo=F}
#### run with R3.4.3
#### ONLY PIECE TO BE CHANGED working directories
wd <- "bootstrap/data/all_landings_2023/"
#output_wd <- "F://D//Expertise//WGNSSK//2016//IC//ple-eche_all_ 2016-4-28 16_46_30_all_all//"
```


```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo=FALSE, warning=FALSE, message=FALSE,fig.height=8)
```


This documet uses Table 2 from CatchAndSampleDataTables. txt from the InterCatch outputs to describe the raising procedures that were made.



```{r}
### libraries
library(data.table)
library(gplots)
library(pander); library(captioner)
library(lattice)
library(ggplot2)
#initialise les fonctions pour légender tables et figures
tabcap <- captioner(prefix = "Table")
figcap <- captioner(prefix = "Figure")

```


```{r}
### read the data
test <- scan(paste(wd,"CatchAndSampleDataTables.txt",sep=""),what='character',sep='\t')
table2 <- test[(which(test=="TABLE 2.")+3):length(test)]
tmp<-table2[-c(1:56)]			  
			  
table2_bis<-data.frame(matrix(tmp,ncol=27,byrow=T), stringsAsFactors =F)
colnames(table2_bis) <- table2[1:27]
table2_bis <- data.table(table2_bis)
table2_bis <- table2_bis[,CATON:=as.numeric(as.character(CATON))]

table2_bis <- table2_bis[,CATON:=CATON/1000]


table2_bis <- table2_bis[,CANUM:=as.numeric(as.character(CANUM))]
table2_bis <- table2_bis[,WECA:=as.numeric(as.character(WECA))]
table2_bis <- table2_bis[,AgeOrLength:=as.numeric(as.character(AgeOrLength))]

table2_bis <- table2_bis[,Area:=as.factor(Area)]
table2_bis <- table2_bis[,Fleet:=factor(Fleet)]
table2_bis <- table2_bis[,Season:=factor(Season)]
table2_bis <- table2_bis[,Country:=factor(Country)]


table2_bis <- table2_bis[,id:=paste(Stock,Country,Area,Season,Fleet)]

table2_bis[Area=="IIIaN                                                       ",'Area'] <- "IIIaN"

table2_bis$CatchCategory <- factor(table2_bis$CatchCategory)

colnames(table2_bis)[colnames(table2_bis)=='CATONRaisedOrImported'] <- 'RaisedOrImported'




		
		
		##modif
		table2_ter <- table2_bis[,-c("Sex","AgeOrLength","CANUM","WECA","LECA")]
		dim(table2_ter)
		table2_ter <- unique(table2_ter)
		dim(table2_ter)
		##modif
		
		SOP <- merge(table2_bis[,list(SOP=sum(CANUM*WECA)/1000000), by=c('CatchCategory')], table2_ter[, list(CATON=sum(CATON)), by=c('CatchCategory')])
		SOP <-SOP[,list(SOP=SOP/CATON), by=c('CatchCategory')]
		
## simple stats on imported data
		
		SampledOrEstimated <- table2_ter[,list(CATON=sum(CATON)), by=c('CatchCategory','RaisedOrImported','SampledOrEstimated')]
		SampledOrEstimated <- SampledOrEstimated[, perc:=round(CATON/sum(CATON)*100), by='CatchCategory']
		SampledOrEstimated <- SampledOrEstimated[order(CatchCategory, perc, decreasing=T),]

		SummaryReport <- SampledOrEstimated[,list(CATON=sum(CATON)), by=c('CatchCategory','RaisedOrImported')]
		SummaryReport <- SummaryReport[, perc:=round(CATON/sum(CATON)*100), by='CatchCategory']
		SummaryReport <- SummaryReport[order(CatchCategory),]		

		
		SampledOrEstimatedArea <- table2_ter[,list(CATON=sum(CATON)), by=c('CatchCategory','RaisedOrImported','SampledOrEstimated','Area')]
		SampledOrEstimatedArea <- SampledOrEstimatedArea[, perc:=round(CATON/sum(CATON)*100), by=c('CatchCategory','Area')]
		SampledOrEstimatedArea <- SampledOrEstimatedArea[order(Area, CatchCategory,perc, decreasing=T),]


		### landings with associated discards
		
	landingsWithAssociatedDiscards <- table2_ter[,list(CATON=sum(CATON)), by=c('id','RaisedOrImported','CatchCategory')]
	tmp <-  landingsWithAssociatedDiscards[(RaisedOrImported=="Imported_Data" & CatchCategory=="Discards")  , id]
	landingsWithAssociatedDiscards <- landingsWithAssociatedDiscards[id%in%tmp & CatchCategory=="Landings", sum(CATON)]/table2_ter[CatchCategory=="Landings",sum(CATON)]*100
	

	
	### compute discards ratio by area for the imported data

	landingsWithAssociatedDiscards2 <- table2_ter[,list(CATON=sum(CATON)), by=c('id','RaisedOrImported','CatchCategory','Area')]
	tmp <-  landingsWithAssociatedDiscards2[(RaisedOrImported=="Imported_Data" & CatchCategory=="Discards")  , id]
	landings <- landingsWithAssociatedDiscards2[id%in%tmp & CatchCategory=="Landings",c("id", "RaisedOrImported","Area", "CATON")]
	colnames(landings)[4] <- "CATON_LAN"
	Discards <- landingsWithAssociatedDiscards2[id%in%tmp & CatchCategory=="Discards",c("id", "RaisedOrImported","Area", "CATON")]
	colnames(Discards)[4] <- "CATON_DIS"
	tmp <- merge(landings, Discards, by=c("id", "RaisedOrImported",      "Area"), all=T)

	DR <- tmp[, list(sumLan=sum(CATON_LAN, na.rm=T), sumDis=sum(CATON_DIS, na.rm=T)), by=list(Area)]
	DR <- DR[,DiscardRatio:=sumDis/(sumDis+sumLan)]
	
		## summary of the catch per gear
		table2_ter <- table2_ter[, Gear:=substr(table2_ter$Fleet,1,3)]
		summCatchPerGear <- table2_ter[, list(sumCatch=sum(CATON)), by=c('Stock','Gear')]
		summCatchPerGear <- summCatchPerGear[, percCatch:=sumCatch/sum(sumCatch)*100, by=c('Stock')]
		summCatchPerGear <- summCatchPerGear[ order(percCatch,decreasing=T),]


		sumLandingsPerGear <- table2_ter[CatchCategory=="Landings",]
		sumLandingsPerGear <- sumLandingsPerGear[, list(sumCatch=sum(CATON)), by=c('Stock','Gear')]
		sumLandingsPerGear <- sumLandingsPerGear[, percCatch:=sumCatch/sum(sumCatch)*100, by=c('Stock')]
		sumLandingsPerGear <- sumLandingsPerGear[ order(percCatch,decreasing=T),]
		
		  
		  
	if(unique(table2_bis$AgeOrLengthType)=="Lngt") AgeTit <- "Length" else AgeTit <- "Age"

### Age Structure

listVAR <- c("CatchCategory",'RaisedOrImported',"SampledOrEstimated","Country","Area","Season","Fleet")

## By Sex

listVAR <- c("CatchCategory",'RaisedOrImported',"SampledOrEstimated","Country","Area","Season","Fleet","Sex")

#table2_bis <- table2_bis[CANUM>0,]

AgeStrucSex <- table2_bis[,meanAge:=weighted.mean(AgeOrLength,CANUM, na.rm=T), by=c(listVAR)]


plotFunction <- function(AgeStrucSexLan,Estim,Sampled){
				plotTitle <- paste("Mean ", AgeTit, " in the ", catchCat, " by ", Var, sep=" ")

	Formula <-paste('meanAge~',Var, sep="")

	boxplot(eval(parse(text=Formula)), data=AgeStrucSexLan,
             boxwex = 0.25, at = which(pos%in%unique(AgeStrucSexLan[,which(colnames(Sampled)==Var)])) + 0.3,
             col = "red",
             main = plotTitle,
             xlab = "",
             ylab = paste("Mean ",AgeTit, sep=""),
             xlim = c(0.5, length(unique(AgeStrucSexLan[,which(colnames(Sampled)==Var)]))+.5), ylim = c(min(AgeStrucSexLan$AgeOrLength), max(AgeStrucSexLan$AgeOrLength)), yaxs = "i", xaxt='n')
	boxplot(eval(parse(text=Formula)), data=Estim, add=TRUE,
             boxwex = 0.25, at = which(pos%in%unique(Estim[,which(colnames(Sampled)==Var)])),
             col = "orange")
	boxplot(eval(parse(text=Formula)), data=Sampled, add=TRUE,
             boxwex = 0.25, at = which(pos%in%unique(Sampled[,which(colnames(Sampled)==Var)])) - 0.3,
             col = "yellow", xaxt='n')
  legend(1, max(AgeStrucSexLan$AgeOrLength), c("Imported", "Raised", "All"),
            fill = c("yellow", "orange", "red"))
	}
	
	
plotFunctionMeanWeight <- function(AgeStrucSexLan=AgeStrucSexLan,Estim=Estim,Sampled=Sampled){
				plotTitle <- paste("Mean ", AgeTit, " in the ", catchCat, " area ", unique(AgeStrucSexLan$Area), sep=" ")

	
AgeStrucSexLan <- data.frame(AgeStrucSexLan)
AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)] <- factor(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])
pos <- levels(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])

Estim <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Estimated_Distribution'),]
Estim[,which(colnames(Estim)==Var)] <- factor(Estim[,which(colnames(Estim)==Var)])

Sampled <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Sampled_Distribution'),]
Sampled[,which(colnames(Sampled)==Var)] <- factor(Sampled[,which(colnames(Sampled)==Var)])
	
	Formula <-paste('WECA~',Var, sep="")

	boxplot(eval(parse(text=Formula)), data=AgeStrucSexLan,
             boxwex = 0.25, at = which(pos%in%unique(AgeStrucSexLan[,which(colnames(Sampled)==Var)])) + 0.3,
             col = "red",
             main = plotTitle,
             xlab = "",
             ylab = paste("Mean Weight", sep=""),
             xlim = c(0.5, length(unique(AgeStrucSexLan[,which(colnames(Sampled)==Var)]))+.5), ylim = c(min(AgeStrucSexLan$WECA), max(AgeStrucSexLan$WECA)), yaxs = "i", xaxt='n')
	boxplot(eval(parse(text=Formula)), data=Estim, add=TRUE,
             boxwex = 0.25, at = which(pos%in%unique(Estim[,which(colnames(Sampled)==Var)])),
             col = "orange")
	
	if(dim(Sampled)[1]>0){
	boxplot(eval(parse(text=Formula)), data=Sampled, add=TRUE,
             boxwex = 0.25, at = which(pos%in%unique(Sampled[,which(colnames(Sampled)==Var)])) - 0.3,
             col = "yellow", xaxt='n')
	}
  legend(1, max(AgeStrucSexLan$WECA), c("Imported", "Raised", "All"),
            fill = c("yellow", "orange", "red"))
}




```
`r tabcap(name="SOP",caption="SOP")`
```{r}
pander(data.frame(SOP))
```
#Raised and imported datas

##Raised discards

In InterCatch, the first step consists in raising the discards volumes for strats with landings and no discards associated. These discards are called in the following table 'Raised_Discards'. The data called 'Imported_Data' are landings or discards volumes imported into InterCatch with or without `r AgeTit` structure.

The proportion of Landings with Discards associated (same strata) is **`r paste(round(landingsWithAssociatedDiscards), 'percent')`**

For these stratas, the discards ratios are the following:

`r pander(data.frame(DR))`

The volumes (and associated proportion) of landings and discards imported  (Imported_Data) or raised (Raised_Discards) are described in the following table.


`r tabcap(name="SummaryReport",caption="Summary of the imported/Raised data")`
```{r}
pander(SummaryReport)
```


##Total catch per gear

The following table gives a summary of the catch (Landings+discards(imported+raised)) per gear (3 first letters of the metier)

`r tabcap(name="SummaryCatch",caption="Total catch (in tonnes) and Percentage of catches per gear")`
```{r}
pander(summCatchPerGear)
```


The following table gives a summary of the landings per gear (3 first letters of the metier)

`r tabcap(name="SummaryLandings",caption="Total landings (in tonnes) and Percentage of catches per gear")`
```{r}
pander(sumLandingsPerGear)
```






## `r AgeTit` distribution

For the imported landings/discards and the raised discards without age distribution, the `r AgeTit` distribution is then computed using the defined allocation scheme. *Sampled_distribution* means that the data (ladings or discards) were input with `r AgeTit` distribution. *Estimated_distribution* means that the inputed/raised valoumes were estimated using the allocation scheme.

In the following tables, CATON is in tonnes

`r tabcap(name="SampledOrEstimated",caption="Summary of the imported/Raised/SampledOrEstimated data")`
```{r}
pander(SampledOrEstimated)
```

`r tabcap(name="SampledOrEstimatedArea",caption="Summary of the imported/Raised/SampledOrEstimated data by area")`
```{r}
pander(SampledOrEstimatedArea)
```
##Impact of the raising on the `r AgeTit` structure

Once the samples imported or raised are identified, it is possible to check the impact of the allocation scheme on the mean age/length of the final age/length distribution of the stock.
The following figures compare the mean age (computed as the weighted mean of the age per strata("CatchCategory",'RaisedOrImported',"SampledOrEstimated","Country","Area","Season","Fleet","Sex")) of the estimated stratas compared to the imported ones and the final distribution. Each individual included in the boxplot corresponds to the weighted mean age of a strata.

###`r  paste("Global mean ",AgeTit)`

```{r meanAge}



AgeStrucSexLan <- AgeStrucSex

ggplot(AgeStrucSexLan, aes(x=CatchCategory, y=meanAge, fill=SampledOrEstimated)) +
  geom_boxplot()				



```
`r tmp<-paste("Mean ",AgeTit, " in the Landings by catch category")`
`r figcap(name="meanAge",caption=tmp)`

###Mean `r AgeTit` per sex

```{r meanAgeSex}
AgeStrucSexLan <- AgeStrucSex[CatchCategory=="Landings",]
				plotTitle <- paste("Mean ", AgeTit, " in the Landings", sep=" ")
				ylabtitle <- paste("mean ",AgeTit)

				
ggplot(AgeStrucSexLan, aes(x=Sex, y=meanAge, fill=SampledOrEstimated)) +
  geom_boxplot()				
				
```
`r tmp<-paste("Mean ",AgeTit, " in the Landings by sex")`
`r figcap(name="meanAgeSex",caption=tmp)`


###Mean `r AgeTit` per area

```{r meanAgeArea}
catchCat <- "Landings"
Var <- "Area"

AgeStrucSexLan <- AgeStrucSex[CatchCategory==catchCat,]
AgeStrucSexLan <- data.frame(AgeStrucSexLan)
AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)] <- factor(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])
pos <- levels(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])

Estim <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Estimated_Distribution'),]
Estim[,which(colnames(Estim)==Var)] <- factor(Estim[,which(colnames(Estim)==Var)])

Sampled <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Sampled_Distribution'),]
Sampled[,which(colnames(Sampled)==Var)] <- factor(Sampled[,which(colnames(Sampled)==Var)])
par(las=2)
plotFunction(AgeStrucSexLan, Estim, Sampled)

```
`r tmp<-paste("Mean ",AgeTit, " in the Landings by Area")`
`r figcap(name="meanAgeArea",caption=tmp)`

###Mean `r AgeTit` per fleet

```{r meanAgeFleet}
catchCat <- "Landings"
Var <- "Fleet"

AgeStrucSexLan <- AgeStrucSex[CatchCategory==catchCat,]
AgeStrucSexLan <- data.frame(AgeStrucSexLan)
AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)] <- factor(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])
pos <- levels(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])

Estim <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Estimated_Distribution'),]
Estim[,which(colnames(Estim)==Var)] <- factor(Estim[,which(colnames(Estim)==Var)])

Sampled <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Sampled_Distribution'),]
Sampled[,which(colnames(Sampled)==Var)] <- factor(Sampled[,which(colnames(Sampled)==Var)])
par(las=2, mar=c(15,4,4,5))
plotFunction(AgeStrucSexLan,Estim,Sampled)

```
`r tmp<-paste("Mean ",AgeTit, " in the Landings by Fleet")`
`r figcap(name="meanAgeFleet",caption=tmp)`

###Resulting `r AgeTit` structure

The following plot shows the percentage of each `r AgeTit` for the sampled strata, estimated and the final `r AgeTit` structure for the landing and discard fractions.


```{r ageStrucTot}
listVAR <- c("CatchCategory","SampledOrEstimated")

AgeStrucSexProp <- AgeStrucSex[,list(sumCANUM=sum(CANUM)), by=c(listVAR,'AgeOrLength')]
AgeStrucSexProp <- AgeStrucSexProp[,propCanum:=sumCANUM/sum(sumCANUM), by=listVAR]

listVAR <- c("CatchCategory")
AgeStrucSexProp2 <- AgeStrucSex[,list(sumCANUM=sum(CANUM)), by=c(listVAR,'AgeOrLength')]
AgeStrucSexProp2 <- AgeStrucSexProp2[,propCanum:=sumCANUM/sum(sumCANUM), by=listVAR]
AgeStrucSexProp2$SampledOrEstimated <- "Final Distribution"

AgeStrucSexProp<- rbind(AgeStrucSexProp,AgeStrucSexProp2)

xlabTitle <- AgeTit
xyplot(propCanum~AgeOrLength|CatchCategory,groups = SampledOrEstimated, data=AgeStrucSexProp,auto.key = list(space = "right", points = TRUE, lines = FALSE), xlab=xlabTitle)

```
`r tmp<-paste(AgeTit, " structure")`
`r figcap(name="ageStrucTot",caption=tmp)`

##Mean weight at `r AgeTit`

the catchAndSampleData also provide the weight at `r AgeTit` per strata for the Sampled/Estimated stratas.
One would also want to check the sampled/estimated and resulting weight at `r AgeTit`. This is produced in the following graph, each boxplot representing the distribution of the weight at `r AgeTit` for the different stratas.


```{r meanWt}
catchCat <- "Landings"
Var <- "AgeOrLength"

AgeStrucSexLan <- AgeStrucSex[CatchCategory==catchCat,]
AgeStrucSexLan <- data.frame(AgeStrucSexLan)
AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)] <- factor(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])
pos <- levels(AgeStrucSexLan[,which(colnames(AgeStrucSexLan)==Var)])

Estim <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Estimated_Distribution'),]
Estim[,which(colnames(Estim)==Var)] <- factor(Estim[,which(colnames(Estim)==Var)])

Sampled <- AgeStrucSexLan[which(AgeStrucSexLan$SampledOrEstimated=='Sampled_Distribution'),]
Sampled[,which(colnames(Sampled)==Var)] <- factor(Sampled[,which(colnames(Sampled)==Var)])
Areas <- unique(AgeStrucSexLan$Area)
#par(las=2, mfrow=c(length(Areas),1 ))
for(i in Areas){
	plotFunctionMeanWeight(AgeStrucSexLan=AgeStrucSexLan[AgeStrucSexLan$Area==i,],Estim=Estim[Estim$Area==i,],Sampled=Sampled[Sampled$Area==i,])
  cat('/r/n/r/n')
	}

```
`r tmp<-paste("Mean ",AgeTit, " in the Landings by Area")`
`r figcap(name="meanWt",caption=tmp)`


The outliers (more than 3 times the standard deviation) are extracted and can be investigated from the following table.


`r tabcap(name="idSamp", cap="Samples that are higher or lower than the average weigth at age/length +/- 3*standard deviation")`
```{r}
catchCat <- "Landings"

AgeStrucSexLan <- data.table(AgeStrucSexLan[AgeStrucSexLan$RaisedOrImported=="Imported_Data" & AgeStrucSexLan$CatchCategory==catchCat,])
AgeStrucSexLan <- AgeStrucSexLan[,AverageWtSize:=mean(WECA), by=list(RaisedOrImported,AgeOrLength,Area)]
AgeStrucSexLan <- AgeStrucSexLan[,stdWECA:=sd(WECA), by=list(RaisedOrImported,AgeOrLength,Area)]
pander(AgeStrucSexLan[WECA>AverageWtSize+3*stdWECA | WECA<AverageWtSize-3*stdWECA   ,c('Country','Fleet','CatchCategory','WECA','AverageWtSize','AgeOrLength','Area'), with=F])
```
